A Bayesian networks approach to infer social changes from burials in northeastern Taiwan during the European colonization period

### 2021-08-24 analysis/paper/paper.Rmd
fix a citation in the rmd and knit

<!-- future work -->
As a large village with continuous occupation since the 14th century in northeastern Taiwan,  the archaeological evidence at Kiwulan provides a special example to explore the reaction of Indigenous people living on the periphery to European colonial presence at a settlement scale that historical records rarely offer. Was its experience with the European presence unique or typical of the region? We need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence between sites on the periphery and colonial centers, such as Heping Dao and Tamsui. The contemporary community of descendants of Kiwulan, the Kavalan people, shows features of a corporate society, such as a matrilineal system with age group administration and ritual practices [@Liu2021]. This indicates that social change has been continuous in this area, from the more centralized network mode of social organization we observed in the burials after European arrival, to the current corporate model of social organization. Future work should focus on collecting data from Indigenous sites related to European contact and Chinese contact for comparison with Kiwulan.

>>>>

<!-- future work -->
As a large village with continuous occupation since the 14th century in northeastern Taiwan,  the archaeological evidence at Kiwulan provides a special example to explore the reaction of Indigenous people living on the periphery to European colonial presence at a settlement scale that historical records rarely offer. Was its experience with the European presence unique or typical of the region? We need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence between sites on the periphery and colonial centers, such as Heping Dao and Tamsui. The contemporary community of descendants of Kiwulan, the Kavalan people, shows features of a corporate society, such as a matrilineal system with age group administration and ritual practices [@LiuPC2021]. This indicates that social change has been continuous in this area, from the more centralized network mode of social organization we observed in the burials after European arrival, to the current corporate model of social organization. Future work should focus on collecting data from Indigenous sites related to European contact and Chinese contact for comparison with Kiwulan.
